Effective Inclusion of Electronic Polarization to a Negatively Charged Membrane

### 2019-03-08 Manuscript/manuscript.tex
Went through the Ca2+ binding section I left a few things undone for Samuli to consider, wheter they would form a good contribution to the work. There still are Figures to be corrected.

The form factors from scattering and C-H bond order parameters from NMR experiments are good measures to validate
the lipid bilayer structure in MD simulations against experiments because they
are related to the bilayer dimensions (area per molecule and thickness) and
conformational fluctuations of individual lipids, respectively, and 
can be directly calculated from MD simulations~\cite{ollila16}.
The X-ray scattering form factors and C-H bond order parameters in
the headgroup and glycerol backbone region of POPS lipid bilayer are already available in the literature \cite{pan14,nmrlipids_proj4}.
Here, we extend the available experimental dataset of POPS lipids by measuring the C--H bond
order parameters also from the acyl chain region using the R-PDLF experiments as described previously \cite{ferreira13}.
We first determined the order parameters of each peak corresponding different C--H bonds detected
in the first dimension (Fig. \ref{R-PDLF}) from the dipolar slices in the indirect dimension (Fig. \ref{R-PDLFslices}).
Because the exact assignment of these peaks was not possible, we assigned the
acyl chain order parameters to mimic the profile predicted by the MD simulations (Fig. \ref{simVSexpNOions_POPS}).
\todo{Tiago has to check and update this and figures in SI.}
\begin{figure}[tb!] 
  \centering 
  \includegraphics[width=\figwidth]{../img/ecc_pops/Order-parameters_form-factors_exp-L17-ECC-lipids.pdf}
  \includegraphics[width=\figwidth]{../img/ecc_pops/pops_chemfig.pdf} 
\hfill
  \caption{\label{simVSexpNOions_POPS} 
    \textbf{Top:} X-ray scattering form factors from simulations with Lipid17/Dang \citep{lipid17-future, dang2006} and 
    the ECC-POPS/ECC-ions \cite{martinek17, Pluhackova2016} compared with experiments~\citep{kucerka14} at 298~K. 
    \textbf{Middle:} Order parameters of POPS head group, glycerol backbone and acyl chains  
    from simulations with the Lipid17 \citep{lipid17-future} and the ECC-POPS models 
    compared with experiments at 298~K. \citep{nmrlipids_proj4}
    Open/closed symbols are used for palmitoyl/oleoyl chains of POPS. 
    \textbf{Bottom:} The chemical structure of POPS and the labeling of the carbon segments. 
  }  
  \todo{J: flip the chemical structure so that it agrees with the left-to-right reading of the plot.} \\
  \todo{The experimental order parameters from sn-1 chain are missing from the figure.} \\
  \todo{We should use the same color coding for the models as in ECC-POPC paper, now it is opposite.} \\
  \todo{Inverted axis and positive value as label would be more clear for the order parameter figure.}
\end{figure}

To evaluate the effect of ECC on the structural properties of POPS lipid bilayer,
we compared the X-ray scattering form factors, C--H bond order parameters and area per molecules
between simulation with original Lipid17 \cite{??} combined with the Dang ion parameters \cite{??},
ECC-POPS simulation and experiments (Fig. \ref{simVSexpNOions_POPS} and table \ref{tab:apls}).
The larger acyl chain order parameters, deviation in the form factor and smaller area per molecule 
in the Lipid17 simulation with Dang ions when compared with experiments indicates that this simulation
predicts too dense lipid bilayer. In previous study \cite{nmrlipids_proj4}, Lipid17 simulated with {\AA}qvist ions~\cite{aqvist90}
gave larger area per molecule (57 \AA$^2$), probably due to the weaker counterion binding affinity.
The Dang ions are used to here to avoid the artificial pairing of ions and condensation at larger salt
concentrations observed with {\AA}qvist \cite{kohagen16, chen07, nmrlipids_proj4} when analysing the effect of ECC. 
The scattering form factors and acyl chain order parameters from ECC-POPS are in good agreement with
experiments (see Fig.~\ref{simVSexpNOions_POPS}), indicating the the bilayer dimensions and acyl chain
conformations are well described by the force field. The area per lipid from ECC-POPS simulation is slightly
smaller than the value reported from SDP model \cite{kucerka14} (table \ref{tab:apls}), but the values
are within each others error bars. 
%In order to obtain the area per lipid from experiments, modeling is used on top of the scattering factors \citep{kucerka14}. 
%In ocntrast, this property is easily extracted directly from simulations with flat phospholipid bilayers. 
%We compare the values from experiments and simulations in Table~\ref{tab:apls}. 
%While the agreement between the scattering form factors from the simulation of a pure POPS bilayer and experiment 
%are excellent (Fig.~\ref{simVSexpNOions_POPS}), there is a non-negligible difference between the values of the area per lipid in Table~\ref{tab:apls}. 
%Since both values are derived from the scattering form factors through modeling of the electron density of the bilayer,
%we cannot decide, which of the values is more reliable.

\begin{table}[tb!] 
\centering
  \caption{Values of the area per lipid (APL) of POPS bilayers 
	   at 298~K with only \ce{Na^+} counterions and no additional ions. 
           Standard deviations are given. \label{tab:apls} } 
  \begin{tabular}{l|c } 
    \multicolumn{2}{c}{POPS} \\
    model          & APL / Å$^2$    \\ 
    \hline 
    Lipid17/Dang              & 53.5$\pm$ 1.7   \\ 
%    Lipid17/ff99              & 57.9$\pm$ 1.7   \\ 
%    \hline 
    ECC-POPS/ECC-ions         & 59.8$\pm$ 1.8   \\ 
%    \hline 
    experiment (SDP model) \citep{kucerka14} & 62.7$\pm$ 1.3  \\ 
%    \hline 
  \end{tabular} \\
  \todo{I removed the value from Lipid17/ff99 simulation because this was not listed in Table I
    and I think that it is more clear if we refer only to the NMRlipids IV result in the text.
    I also changed the experimental value to correspond the value in the cited paper and added the error bar.}
\end{table}

@@

@@

Despite the success in simulating bilayer dimensions and acyl chain conformations,
lipid force fields have typically problems in capturing the correct glycerol backbone and
headgroup order parameters and conformations \cite{botan15,ollila16,nmrlipids_proj4}.
This is the case also for Lipid17 POPS simulations, where the headgroup and glycerol backbone
order parameters are quite far from experimental values with both Dang (Fig. \ref{simVSexpNOions_POPS})
and {\AA}qvist \cite{nmrlipids_proj4} ions. The values from ECC-POPS simulation are closer,
but not in full agreement with experiments (Fig. \ref{simVSexpNOions_POPS}).
Especially, the experimentally observed more negative values of the $\beta$-carbon
and the larger forking (difference of order parameters between hydrogens attached to the
same carbon) of the $\alpha$-carbon in POPS than in POPC \cite{nmrlipids_proj4}
are not reproduced by neither of the models here. These differences are previously related
to a more rigid structure of PS headgroup and the stiffness of the C$_\alpha$-C$_\beta$-C$_\gamma$-O$_\gamma$
dihedral, which is \todo{the dihedral should be plotted and sentence finished.}

%The head group order parameters $\alpha$ and $\beta$ are highly relevant for this work,
%as they are being used in the electrometer concept \cite{altenbach84, catte16, melcr18}.
%For POPC in pure water, the order parameter $\beta$ agrees well with the experiment, while the order parameter $\alpha$ is somewhat lower.

In conclusion, the ECC-POPS model reproduces the lipid bilayer dimensions and
acyl chain conformations in good agreement with experiments, but has room for
improvement in the glycerol backbone and headgroup region. Therefore,
its structural accuracy is in the same level as in other state-of-the-art
lipid force fields \cite{botan15, ollila16, Pluhackova2016, nmrlipids_proj4}.

>>>>

\begin{figure}[tb!] 
  \centering 
  \includegraphics[width=\figwidth]{../img/ecc_pops/Order-parameters_form-factors_exp-L17-ECC-lipids.pdf}
  \includegraphics[width=\figwidth]{../img/ecc_pops/pops_chemfig.pdf} 
\hfill
  \caption{\label{simVSexpNOions_POPS} 
    \textbf{Top:} X-ray scattering form factors from simulations with Lipid17/Dang \citep{lipid17-future, dang2006} and 
    the ECC-POPS/ECC-ions \cite{martinek17, Pluhackova2016} compared with experiments~\citep{kucerka14} at 298~K. 
    \textbf{Middle:} Order parameters of POPS head group, glycerol backbone and acyl chains  
    from simulations with the Lipid17 \citep{lipid17-future} and the ECC-POPS models 
    compared with experiments at 298~K. \citep{nmrlipids_proj4}
    Open/closed symbols are used for palmitoyl/oleoyl chains of POPS. 
    \textbf{Bottom:} The chemical structure of POPS and the labeling of the carbon segments. 
  }  
\todo{J: flip the chemical structure so that it agrees with the left-to-right reading of the plot.}
\end{figure}

\begin{table}[tb!] 
\centering
  \caption{Values of the area per lipid (APL) of POPS bilayers 
	   at 298~K with only \ce{Na^+} counterions and no additional ions. 
           Standard deviations are given. \label{tab:apls} } 
  \begin{tabular}{l|c } 
    \multicolumn{2}{c}{POPS} \\
    model          & APL / Å$^2$    \\ 
    \hline 
    Lipid17/Dang              & 53.5$\pm$ 1.7   \\ 
    Lipid17/ff99              & 57.9$\pm$ 1.7   \\ 
    \hline 
    ECC-POPS/ECC-ions         & 59.8$\pm$ 1.8   \\ 
    \hline 
    experiment (SDP model) \citep{kucerka14} & 62.3  \\ 
    \hline 
  \end{tabular} 
\end{table}

We compared X-ray scattering form factors and NMR order parameters of bilayers
in pure water without any ions (or only counter ions)
from simulations and experiments. 
The experimental X-ray scattering form factors 
of a bilayer are well reproduced by ECC-POPS, 
while for Lipid17/Dang they are in a slight disagreement
(see Fig.~\ref{simVSexpNOions_POPS}). 
The agreement of Lipid17 can be improved by using different parametes for ions \cite{aqvist90},
which, however, bring other kinds of artifacts with them 
(e.g., incorrect pairing of ions, condensation at larger salt concentrations \cite{kohagen16, chen07, nmrlipids_proj4}).

%The area per lipid is often used as a relatively simple structural parameter reporting on the bilayer properties and the packing of lipids. 
In order to obtain the area per lipid from experiments, modeling is used on top of the scattering factors \citep{kucerka14}. 
In ocntrast, this property is easily extracted directly from simulations with flat phospholipid bilayers. 
We compare the values from experiments and simulations in Table~\ref{tab:apls}.

@@

While the agreement between the scattering form factors 
from the simulation of a pure POPS bilayer and experiment 
are excellent (Fig.~\ref{simVSexpNOions_POPS}),
there is a non-negligible difference between the values of the area per lipid in Table~\ref{tab:apls}. 
Since both values are derived from the scattering form factors through modeling of the electron density of the bilayer,
we cannot decide, which of the values is more reliable. 
In general, we can conclude that ECC-lipids
%the presented lipid models with ECC-lipids 
reproduce the experimental structural parameters of the lipid bilayers 
with a comparable accuracy to existing state-of-the-art lipid models~\citep{botan15, ollila16, Pluhackova2016}.

@@

The head group and acyl chain order parameters within ECC-lipids
are in general in a good agreement with the experimental values 
as shown in Fig. \ref{simVSexpNOions_POPS}. 
The acyl chain order parameters in particular agree with the experimental data within error bars.
The order parameters of the head groups are at an accuracy comparable to 
other currently available classical models of lipids \citep{botan15, catte16, Pluhackova2016, nmrlipids_proj4}. 
In particular, ECC-POPS is closer to the experimental order parameters 
than Lipid17 with either Dang \cite{smith94,chang1999,dang2006} 
or ff99 ions \citet{aqvist90}.

%The head group order parameters $\alpha$ and $\beta$ are highly relevant for this work,
%as they are being used in the electrometer concept \cite{altenbach84, catte16, melcr18}.
%For POPC in pure water, the order parameter $\beta$ agrees well with the experiment, while the order parameter $\alpha$ is somewhat lower. 
The head group order parameters of PS are a bit more complicated 
compared to PC as the order parameter $\alpha$ exhibits a notable forking 
(i.e., the hydrogen atoms in the group give rise to different order parameters, see Fig.~\ref{simVSexpNOions_POPS}).
One of the order parameters of ECC-POPS, $\alpha_1$, agrees well with the experiment, 
while the other, $\alpha_2$, adopts a higher value underestimating the experimentally reported forking. 
There is only one order parameter $\beta$ in POPS, 
which has a higher value closer to zero in the ECC-lipids model than in experiment. 
Such a feature suggests that the model likely overestimates the orientational freedom of the PS head group.
